Supplemental Materials: \emph{Simultaneous segmentation and registration of diffusion MR images of the brain driven by active-contours}

### 2015-03-24 Publications/2015-NeuroImage/supplemental.tex
editions after @daducci review
Updates after Yasser and Dominique

% @Author: Oscar Esteban
% @Date:   2015-02-20 15:40:09
% @Last Modified by:   Oscar Esteban
% @Last Modified time: 2015-03-18 15:12:35

@@

\paragraph{Obtaining the update equation} Once the shape gradients are obtained \eqref{eq:gradient_final},
  we apply the Fourier Transform to solve the \gls*{pde}:
\begin{align}
  \frac{\partial E(\vec{u})}{\partial \vec{u}_k} \equiv \frac{\vec{u}_k^{t+1} - \vec{u}_k^t}{\delta} \implies
  \vec{u}_k^{t+1} &= \delta \Big( \vec{g}_k + \boldsymbol{\alpha} \vec{u}_k
  - \boldsymbol{\beta} \Delta \vec{u}_k \Big) + \vec{u}_k^t, \notag \\
  \vec{u}_k^{t+1} &= \delta \vec{g}_k +  \boldsymbol{\alpha} \delta \vec{u}_k
  - \boldsymbol{\beta} \delta \Delta \vec{u}_k + \vec{u}_k^t
  \label{eq:discrete_derivative}
\end{align}

@@

\paragraph{Efficient interpolation using sparse matrices}
During the registration process, every iteration requires computing the product of all the gradients
  $\vec{g}_k$ associated to the control point $k$, and computed at the current position
  $\vec{v}_i$ by the corresponding weights $\psi_{ik} = \psi_k(\mathbf{v}_i)$ of interpolating functions
  \eqref{eq:shape_gradient_final}.
In order to optimize multiplications and summations, all the $\psi_{ik}$ are collected in a
  matrix $\boldsymbol{\Psi} = (\psi_{ik})$.
Given the limited support of the basis function $\psi$, $\boldsymbol{\Psi}$ will hold the property
  of being sparse, as only few $\psi_{ik} > 0$ in the surroundings of $\vec{v}_i$.
Then, the gradients $g_k$ are easily computed using the matrix product:
\begin{equation}
  \big(\vec{g}_k \big) = \boldsymbol{\Psi} \cdot \big( w_i \, \bar{s}_i \big)^T
  \label{eq:sparse_matrix}
\end{equation}

>>>>

% @Author: Oscar Esteban
% @Date:   2015-02-20 15:40:09
% @Last Modified by:   Oscar Esteban
% @Last Modified time: 2015-03-24 16:12:10

@@

\paragraph{Obtaining the update equation} To solve the differential equation in \eqref{eq:gradient_descent},
  we use a semi-implicit Euler scheme, where the shape-gradients $\vec{g}_k$ are explicit:
\begin{align}
  \vec{g}_k^{t+1} &= \vec{g}_k^{t} - \vec{u}_k^{t}, \notag \\
  \vec{u}_k^{t+1} &= \vec{u}_k^{t} + \delta \, \Big( \vec{g}_k^{t+1} + (\boldsymbol{\alpha} - \boldsymbol{\beta} \Delta) \, \vec{u}_k^{t+1} \Big),
  \label{eq:semiimplicit_euler}
\end{align}
and introducing the $\vec{g}_k^{t+1}$ in the second equation:
\begin{align}
  \vec{u}_k^{t+1} &= \vec{u}_k^{t} + \delta \Big( \vec{g}_k^{t} + \delta \vec{u}_k^t + (\boldsymbol{\alpha} - \boldsymbol{\beta} \Delta) \, \vec{u}_k^{t+1} \Big) \notag \\
  \vec{u}_k^{t+1} &= \frac{ (1+\delta^2) \, \vec{u}_k^{t} + \vec{g}_k^{t}} {1 - \boldsymbol{\alpha} + \boldsymbol{\beta} \Delta}
  \label{eq:discrete_derivative}
\end{align}

@@

\paragraph{Efficient interpolation using sparse matrices}
During the registration process, every iteration requires computing the product of all the gradients
  $\vec{g}_k$ associated to the control point $k$, and computed at the current position
  $\vec{v}_i$ by the corresponding weights $\psi_{ik} = \psi_k(\mathbf{v}_i)$ of interpolating functions
  \eqref{eq:shape_gradient_final}.
In order to optimize multiplications and summations, all the $\psi_{ik}$ are collected in a
  matrix $\boldsymbol{\Psi} = (\psi_{ik})$.
Given the limited support of the basis function $\psi$, $\boldsymbol{\Psi}$ will hold the property
  of being sparse, as only few $\psi_{ik} > 0$ in the surroundings of $\vec{v}_i$.
Then, the gradients $\vec{g}_k$ are easily computed using the matrix product:
\begin{equation}
  \big(\vec{g}_k \big) = \boldsymbol{\Psi} \cdot \big( w_i \, \bar{s}_i \big)^T
  \label{eq:sparse_matrix}
\end{equation}
